---
title: "Wearable Device Prediction"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

## Reading and Pruning the Data

We read in the data and remove columns which are overwhelmingly missing. We also remove the index column and the cvtd_timestamp since raw timestamp should suffice.

```{r}
library(caret)
traindata<- read.csv("pml-training.csv", na.strings="")
traindata<- traindata[,c(2:4,5:11,37:49,60:68,84:86,102,113:124,140,151:160)]
spl<- split(traindata,traindata$user_name)
```
## Fitting a Random Forest

We fit a random forest to the training data. For this purpose, we split the data into 6 sets, based on the `user_name`, in order to save on RAM and with the reasoning that each user has their own distinct behavior patterns.

```{r}
spltrain <- split(traindata,traindata$user_name)
fitlst<- NULL
for (i in 1:6) {
  curfit<-train(classe~ ., data=spl[[i]],method="rf",prox=TRUE)
  fitlst<- c(fitlst,list(curfit))
}

```

Now we load the test data, split it into groups by `user_name`, predict using the appropriate model, and combine the results.

```{r}
library(dplyr)
testdata<- read.csv("pml-testing.csv")
spltest <- split(testdata,testdata$user_name)
predtest<- NULL
for (i in 1:6) {
  spltest[[i]]$pred<-predict(fitlst[[i]],spltest[[i]])
  predtest<-rbind(predtest,spltest[[i]])
}
predtest<- select(predtest,problem_id,pred)
predtest<- arrange(predtest,problem_id)
predtest
```

##Notes on cross-validation and out-of-sample error

We use caret's built-in model selection, which uses re-sampling (bootstrapped) with 25 repetitions. We do this separately for each `user_name`. Caret's built-in training function also uses resampling to calculate the accuracy. Because we use resampling, the corresponding estimate of the error rate is likely smaller than the true out-of-sample error rate.

Below, we calculate our approximation of the out-of-sample error rate by averaging across all `user_name`s.

```{r}
totrate <- 0
for (i in 1:6){
  totrate <- totrate + 1- max(fitlst[[i]]$results$Accuracy)
}
meanrate<- totrate/6
meanrate
```

This, our estimation of the error rate is `r meanrate`

